# Emissions {#emissions}

**This chapter will show you how to create charts** showing annual emissions for one or more *categories* and one or more *pollutants*.

It will also show you how to customize those charts with:
    
- a unique *color* for each category;
- directly labeled *values* at specific years; and
- reader-friendly *names* for categories.

Finally, it will show you how to create *area charts*, consistent with all of the above.

All of what you learn will be extensible to the chapters on [Throughputs], [Growth Profiles], [Emission Factors], and [Control Factors].

## Key functions

This section relies on three key functions:

- `chart_annual_emissions_by()`
- `filter_categories()`
- `filter_pollutants()`

The `chart_annual_emissions_by()` function --- and its shorter form, `chart_annual_emissions()` --- will be the foundation of this section. To learn more about `chart_annual_emissions_by()`, you can always type `help("chart_annual_emissions_by")`.

If you have read the vignette *Introduction to BY Data*^[This is being migrated from an R vignette to a GitBook like this one. Stay tuned!] then you are already familiar with `filter_categories()` and `filter_pollutants()`.

## Charting annual emissions

Let's try it out on a relatively simple case: charting NOx emissions from BY2011 category `#283`, natural gas consumed for residential water heating.

```{r BY2011_annual_emission_data-283-NOx, fig.height=3, out.height=300}
#
# This is an absolutely minimal example for `chart_annual_emissions()`.
# We're going to build on it, adding improvements as we go.
#
BY2011_annual_emission_data %>%
  filter_categories(
    283) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_emissions()
```

That's a great first step!

## Coloring by category

If you want to chart more than one category, you'll probably want a way to distinguish them visually. Let's use `color`. 

```{r BY2011_annual_emission_data-283-284-NOx, fig.height=3, out.height=300}
#
# Here we'll color by `category`. You could also color by `cnty_abbr`. (Try it!)
#
# By default, `filter_categories()` adds a new column `category`. If you don't
# name the categories --- which we will do, further along --- then they are 
# simply prefixed with "#". So, for example, when `cat_id` is 283, the default 
# for `category` will be "#283".
#
# The suffix `_by()` --- on `chart_annual_emissions_by()` --- is mostly just
# there for readability. Things should still work if you forget and use
# `chart_annual_emissions()` instead.
# 
BY2011_annual_emission_data %>%
  filter_categories(
    283, 
    284) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_emissions_by(
    color = category)
```

The default colors are provided by `scale_color_tableau()`. You can try using a different color scheme if you like. Go to the "Help" pane in RStudio, and type "scale_color_", then wait. You should see a list of choices pop up.

```{block type="note"}
If you wind up with more than ten colors, you may need to rethink the chart. Maybe grouping categories would be helpful? In later sections, we will learn how to easily do this.
```

## Flagging specific values

Displaying exact values on a chart can be helpful to the reader. Then, they don't have to consult a separate table. 

To do this, simply supply `flag_years`. If you want to change the template for what exactly is displayed --- digits, text, etc. --- you can supply `flag_labels` too. See the [Appendix](#appendix-howto-flag_labels).

```{r BY2011_annual_emission_data-283-284-NOx-flag_years, fig.height=3, out.height=300}
#
# Use `flag_years` to display values directly in the chart.
#
BY2011_annual_emission_data %>%
  filter_categories(
    283, 
    284) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_emissions_by(
    color = category,
    flag_years = CY(1993, 2011, 2030))
```

## Humanizing category names

You can also, in the process of filtering, supply more human-readable names for the categories.

```{r BY2011_annual_emission_data-283-284-NOx-humanize_category_names, fig.height=3, out.height=300}
#
# Names supplied to `filter_categories()` appear in the data as a new column,
# `category`. Following that, we can ask `chart_annual_emissions_by()` to use
# `category` as the basis for coloring.
# 
BY2011_annual_emission_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_emissions_by(
    color = category,
    flag_years = CY(1993, 2011, 2030))
```

## Titling and captioning

Last but not least, you can provide a `title`, `subtitle`, and/or `caption`.

```{r BY2011_annual_emission_data-chart_annual_emissions_by-category-with_title}
#
# An explicit `title`, `subtitle`, and/or `caption` can be supplied to
# `chart_annual_emissions_by()`. This is strongly encouraged!!
#
# The default caption is just "DRAFT YYYY-mm-dd."
# 
BY2011_annual_emission_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_emissions_by(
    color = category,
    flag_years = CY(1993, 2011, 2030),
    title = "Residential NG Combustion: Space and Water Heating",
    subtitle = "Between CY1993-2002, controls on NOx from water heating were strengthened.")
```

## Making an area chart

Stacking can be a good way to show how the total and the proportions vary over time. (Compare with the line-charts above, which are not stacked.)

To create a stacked ("area") chart, simply pass `fill = ...` instead of `color = ...`.

```{r BY2011_annual_emission_data-283-284-NOx-fill_by_category}
#
# `fill` can be supplied in place of `color`.
#
BY2011_annual_emission_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_emissions_by(
    fill = category,
    title = "Residential NG Combustion: Space and Water Heating",
    subtitle = "Combined emissions from water and space heating have declined, but are projected to rise reach 1990 levels by 2030.")
```

## More than one pollutant

Above, we've kept things simple by just focusing on `NOx`. But you'll often want to show more than one pollutant at the same time. You can definitely do that!

```{r BY2011_annual_emission_data-283-284-multiple_pollutants, fig.height=6, out.height=600}
#
# If multiple pollutants are represented in the data that is passed to
# `chart_annual_emissions_by()`, it will try to be helpful by creating a
# sub-plot for each pollutant. This ought to be a safe approach.
# 
BY2011_annual_emission_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  filter_pollutants(
    "NOx",
    "PM10") %>%
  chart_annual_emissions_by(
    color = category,
    title = "Residential NG Combustion: Space and Water Heating",
    subtitle = "Between CY1993-2002, controls on NOx from water heating were strengthened.")
```
