# Throughputs {#throughputs}

**This chapter is very similar to [the previous chapter ("Emissions")][Emissions].** If you haven't read that one yet, it would probably be a good idea to read it first!

The first section in this chapter, [Obtaining throughput data] (below), is a bit of a digression. If you like, you can just skip ahead to [Charting annual throughputs].

## Obtaining throughput data

In the [Emissions] chapter, we had a nice "published" dataset ready at hand: `BY2011_annual_emission_data`.

Unfortunately, there are no analogously "published" area-source throughputs for `BY2011`. However, we can fairly easily (re-)produce CY1990-2040 area-source estimates from the information contained in the `BY2011` DataBank t-tables. Then, we can extract throughputs from those. That's what we'll do in the subsections below. Again, this is a bit of a digression --- you can just skip ahead to [Charting annual throughputs] and come back to this later, if you like.

### Recreating BY2011 area-source projections

In the chunk of code below, we recreate the entire BY2011 area-source inventory, from CY1990 through CY2040. (It's actually pretty fast.)

```{r BY2011_area_source_projection_data, cache=TRUE}
#
# Step 1: `BY_area_source_projections()` recreates an entire inventory. For
# each category, it assembles `tput_qty`, `ef_qty`, and `cf_qty`, and then
# calculates `ems_qty` from those.
#
# It can also do much more --- see `help("BY_area_source_projections")` to learn
# how you can "swap in" alternative throughputs, emission factors, etc.
#
# **NOTE:** For BY2011, in some cases, the calculated emission estimates may
# differ from the published emission estimates. See the Appendix for an
# explanation and a specific example.
# 
BY2011_area_source_projection_data <-
  BY(2011) %>%
  BY_area_source_projections(
    verbose = TRUE)

head(BY2011_area_source_projection_data)
```

The resulting `BY2011_area_source_projection_data` now contains columns for: 

- `ems_qty` and `ems_unit` (emissions);
- `ef_qty` (emission factors);
- `cf_qty` (uncontrolled fractions); and
- `tput_qty` and `tput_unit` (throughputs). 

### De-duplicating recreated BY2011 throughputs

Next, we derive `BY2011_area_source_throughput_data` from `BY2011_area_source_projection_data`.

The fact that `BY2011_area_source_projection_data` contains `tput_qty` and `tput_unit` makes it *almost* suitable for use with `chart_annual_throughputs_by()`. But, `BY2011_area_source_projection_data` is in long ("tidy") format, suitable for use with `chart_annual_emissions_by()`. This means that the values in the `tput_qty` column are repeated once per pollutant. If we just added them up, we'd end up with a big over-estimate. 

We'll use `distinct()` to solve the problem.

```{r BY2011_area_source_throughput_data-distinct, cache=TRUE}
#
# We can use `distinct()` to de-duplicate throughputs.
# 
BY2011_area_source_throughput_data <-
  BY2011_area_source_projection_data %>%
  distinct(
    year, 
    cat_id,
    cnty_abbr,
    tput_qty,
    tput_unit)
```

Now we have a clean `BY2011_area_source_throughput_data`.

Again, if you're reading in throughput data from a .CSV or .XLSX file, you won't have to worry at all about the steps above. Just skip ahead to [the next section][Charting annual throughputs]!

## Charting annual throughputs

Let's examine a category where the growth in throughput varies by county over time. A good example is `#761` "Sanitary Sewers".

```{r BY2011_area_source_projection_data-283-284-chart_annual_throughputs_by, cache=TRUE, dependson="BY2011_area_source_projection_data"}
#
# Even though the relevant throughputs appear twice in these data --- once for
# `ROG` rows, and once for `CH4` --- `chart_annual_throughputs_by()` tries not
# to double-count. See `help("chart_annual_throughputs_by")` for details.
# 
BY2011_area_source_projection_data %>%
  filter_categories(
    "Sanitary Sewers" = 761) %>%
  chart_annual_throughputs_by(
    color = cnty_abbr,
    title = "Sanitary Sewers",
    subtitle = str_c(
      "Sanitary sewers (category #761) tracks with household population.",
      "The relevant growth profile (#657, in DataBank) also varies by county.",
      sep = "\n"))
```

As you can see, the syntax for `chart_annual_throughputs_by()` is identical to that for `chart_annual_emissions_by()`. If you used `chart_annual_emissions()` and didn't supply `color = cnty_abbr`, you would see the regional totals instead.

```{block type="note"}
Above, we've colored by `cnty_abbr`, rather than `cat_id`. This often makes sense when working with throughput data, since different categories often are estimated in terms of different throughput units. So, displaying throughputs for different categories on the same scale may not always make sense (although when the units are the same, it does).
```

What if we wanted to understand these throughputs in terms of percent change (%), rather than absolute units ("1000 Pounds")? Let's move on to the next two chapters, [Relative Changes] and [Growth Profiles].
