# Relative Changes

This chapter shows how to chart *relative changes* in emissions and/or throughputs. By "relative change" we simply mean the percent change relative to a given year (i.e., the "base year" or "reference year").

We don't actually need to obtain any growth profile data yet; the function `chart_annual_growth_by()` will simply normalize whatever we supply it with. In the next chapter, we'll look at [growth profiles][Growth Profiles] more specifically.

## Relative changes in emissions

We can easily examine relative growth using `chart_annual_growth_by()`. It expects the same input as `chart_annual_emissions()`; the only additional requirement is that you supply a `base_year`. 

Here is an example. It's exactly the same as we saw in the [Emissions] chapter, except that we're swapping `chart_annual_growth_by()` in for `chart_annual_emissions_by()`.

```{r BY2011_annual_emission_data-283-284-NOx-chart_annual_growth_by, cache=TRUE}
#
# The `base_year` can be prefixed with `CY`, `BY`, `RY` or `PY`;
# `chart_annual_growth_by()` will ignore the differences.
# 
BY2011_annual_emission_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_growth_by(
    color = category,
    base_year = CY(2011)) # required; can be BY or CY
```

We can enhance it in exactly the same ways that we learned in the [Emissions] chapter, too. Using `flag_years` can be especially helpful, because then we can read off the exact growth for a given future (or historical) year.

```{r BY2011_annual_emission_data-283-284-NOx-chart_annual_growth_by-enhanced, cache=TRUE}
BY2011_annual_emission_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_growth_by(
    color = category,
    base_year = BY(2011),
    flag_years = CY(1993, 2002, 2030),
    title = "Residential NG Combustion: Space and Water Heating",
    subtitle = str_c(
      "NOx emissions from water and space heating are projected to grow by +24% from CY2011 to CY2030.",
      "Labeled values are percent changes relative to the base year (CY2011).",
      sep = "\n")) # combine lines with a newline
```

## Relative changes in throughputs

You can reuse the chunk above to work with throughput data, provided you drop the `filter_pollutants(...)` clause (two lines).

```{r BY2011_area_source_throughput_data-283-284-chart_annual_growth_by-cat_id, cache=TRUE}
BY2011_area_source_throughput_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  chart_annual_growth_by(
    color = category,
    base_year = BY(2011),
    flag_years = CY(1993, 2002, 2030),
    title = "Residential NG Combustion: Space and Water Heating",
    subtitle = str_c(
      "Natural gas consumed by water and space heating is projected to grow by +24% from CY2011 to CY2030.",
      "Labeled values are percent changes relative to the base year (CY2011).",
      sep = "\n")) # combine lines with a newline
```

We can also look at the trend in total throughput for these two caegories, instead of splitting them apart. Just omit the `color = cat_id`, and swap in `chart_annual_growth()` (without the suffix `_by()`).

```{r BY2011_area_source_throughput_data-283-284-chart_annual_growth, cache=TRUE}
BY2011_area_source_throughput_data %>%
  filter_categories(
    "#283 Space Heating" = 283, 
    "#284 Water Heating" = 284) %>%
  chart_annual_growth(
    base_year = BY(2011),
    flag_years = CY(1993, 2002, 2030),
    title = "Residential NG Combustion: Space and Water Heating",
    subtitle = str_c(
      "Natural gas consumed by water and space heating is projected to grow by +24% from CY2011 to CY2030.",
      "Labeled values are percent changes relative to the base year (CY2011).",
      sep = "\n")) # combine lines with a newline
```


```{block type="note"}
If you are trying this, you might notice that the throughput forecast doesn't follow the same trajectory as the NOx forecast. See the [Appendix](#appendix-case-cat_283) for an explanation and an advanced exercise.
```
