# Emission Factors

For some categories, emission factors vary over time. A good category to demonstrate this with is `#749` "Agricultural Land Preparation". 

Here, `annualize_DB_emission_factors()` does the work of carrying emission factors forward in time. This means that for each value of `year`, the "most recent" emission factor, as of that year, is stored in `ef_qty`. 

```{r BY2011_ag_land_prep_ef_data}
BY2011_ag_land_prep_ef_data <-
  BY(2011) %>%
  DB_area_source_emission_factors(
    verbose = TRUE) %>%
  filter_categories(
    "#749 Agricultural Land Prep" = 749) %>%
  annualize_DB_emission_factors(
    years = CY(1990:2030),
    verbose = TRUE)
```

In practice, `chart_annual_emission_factors_by()` is almost exactly identical to `chart_annual_emissions_by()`. The only real difference is in the default formatting of the values displayed directly in the chart by `flag_years`.

```{r BY2011_ag_land_prep_ef_data-chart_annual_emission_factors}
BY2011_ag_land_prep_ef_data %>%
  chart_annual_emission_factors(
    flag_years = CY(2011),
    title = "Agricultural Land Preparation",
    subtitle = "Emission factors pulled from DataBank (t1326).") 
```

```{block type="note"}
In DataBank, emission factors are stored in units of "pounds per unit of throughput". If you change the value of `ef_unit` (and alter the values in `ef_qty` accordingly), that will be reflected in the output of `chart_annual_emission_factors()`.
```