# Third-Party Data

You can use the `chart_*_by()` family of functions with third-party data too.

This set of examples pulls from an R library, `ECDMS`, that contains tidied data harvested from the [California Energy Consumption Database (ECDMS)](http://www.ecdms.energy.ca.gov). These data, published by the California Energy Commission (CEC), provide estimates of annual natural gas and electricity consumption in California.

To gain access, we just type `library(ECDMS)`.

```{r library-ECDMS, cache = FALSE}
library(inventory)
library(ECDMS)
```

## Natural Gas Consumption

In the `ECDMS_gas_county_data` dataset, natural gas consumption data are split by county, sector, and year. This dataset covers all of California, from `r first(sort(pull_distinct(ECDMS_gas_county_data, year)))` through `r last(sort(pull_distinct(ECDMS_gas_county_data, year)))`. Here's a preview:

```{r ECDMS_gas_county_data}
head(ECDMS_gas_county_data)
```

```{block type="note"}
What we're interested in is `tput_qty` --- so `chart_annual_throughputs_by()` will "just work".
```

### San Francisco Bay Area (9 Counties)

Let's take `SFBA_gas_county_data` to be the subset belonging to `r str_and(names(DST_COUNTY_NAMES))` counties.

```{r SFBA_gas_county_data}
#
# Note: `DST_COUNTY_NAMES` is supplied by the `inventory` package.
#
SFBA_gas_county_data <-
  ECDMS_gas_county_data %>%
  filter(
    county %in% names(DST_COUNTY_NAMES)) 
```

Now let's create a set of plots:

- Throughput, total
- Relative growth, total
- Stacked throughputs, by sector
- Relative growth, by sector
- Stacked throughputs, by county
- Relative growth, by county

For the relative-growth charts, we'll specify `base_year = CY(2011)`.

#### SFBA: Total Consumption {-}

```{r SFBA_gas_county_data-MMthm-chart}
SFBA_gas_county_data %>%
  chart_annual_throughputs(
    flag_years = CY(1990, 2011))
```

#### SFBA: Total Growth {-}

```{r SFBA_gas_county_data-growth-chart}
SFBA_gas_county_data %>%
  chart_annual_growth(
    base_year = CY(2011),
    flag_years = CY(1990))
```

#### SFBA: Consumption by Sector {-}

```{r SFBA_gas_county_data-MMthm-sector-chart}
SFBA_gas_county_data %>%
  chart_annual_throughputs_by(
    fill = sector)
```

#### SFBA: Growth by Sector {-}

```{r SFBA_gas_county_data-growth-sector-chart}
SFBA_gas_county_data %>%
  chart_annual_growth_by(
    color = sector,
    base_year = CY(2011),
    flag_years = CY(1990))
```

Here we see that, in the nine-county Bay Area, Non-Residential gas consumption has been trending slightly upward over time, at about 0.3% growth per year. Residential consumption, on the other hand, has been trending slightly downward, at about -0.3% per year.

#### SFBA: Consumption by County {-}

```{r SFBA_gas_county_data-chart-MMthm-county-chart}
SFBA_gas_county_data %>%
  chart_annual_throughputs_by(
    fill = county)
```

#### SFBA: Growth by County {-}

```{r SFBA_gas_county_data-growth-county-chart}
SFBA_gas_county_data %>%
  chart_annual_growth_by(
    color = county,
    base_year = CY(2011))
```

***

### PG&E Planning Area

Above, we saw that `ECDMS_gas_county_data` is split by `county`. There is a different tabular dataset, `ECDMS_gas_plan_data`, that is split instead by `plan` ("planning area").

Let's take `PGE_gas_plan_data` to be the subset where `plan` is "Pacific Gas and Electric".

```{r PGE_gas_plan_data}
PGE_gas_plan_data <-
  ECDMS_gas_plan_data %>%
  filter(
    plan == "Pacific Gas and Electric") 
```

Now let's recreate a similar set of plots:

- Throughput, total
- Relative growth, total
- Stacked throughputs, by sector
- Relative growth, by sector

```{block type="note"}
The total annual consumption is higher in `PGE_gas_plan_data` --- more like 5 billion therms, versus the 3 billion therms we saw in `SFBA_gas_county_data`.
```

#### PG&E: Total Consumption {-}

```{r PGE_gas_plan_data-MMthm-chart}
PGE_gas_plan_data %>%
  chart_annual_throughputs(
    flag_years = CY(1990, 2011))
```

#### PG&E: Total Growth {-}

```{r PGE_gas_plan_data-growth-chart}
PGE_gas_plan_data %>%
  chart_annual_growth(
    base_year = CY(2011),
    flag_years = CY(1990))
```

#### PG&E: Consumption by Sector {-}

```{r PGE_gas_plan_data-MMthm-sector-chart}
PGE_gas_plan_data %>%
  chart_annual_throughputs_by(
    fill = sector)
```

#### PG&E: Growth by Sector {-}

```{r PGE_gas_plan_data-growth-sector-chart}
PGE_gas_plan_data %>%
  chart_annual_growth_by(
    color = sector,
    base_year = CY(2011),
    flag_years = CY(1990))
```

If we like, we can collapse these sectors into "Residential", "Industrial/Commercial", and "Other".

```{r PGE_gas_plan_data-growth-sector-regroup-chart}
PGE_gas_plan_data %>%
  mutate_at(
    vars(sector),
    ~ fct_collapse(
      .,
      "Residential" = "Residential",
      "Commercial/Industrial" = c(
        "Industry", 
        "Commercial Building",
        "Commercial Other"),
      group_other = TRUE)) %>%
  chart_annual_growth_by(
    color = sector,
    base_year = CY(2011),
    flag_years = CY(1990))
```

This shows something different than the analysis by county. 

- **Residential consumption (MMthm).** In this `PGE_gas_plan_data`, we see about 2 billion thm/yr. In the `SFBA_gas_county_data`, we only saw about 1 billion thm/yr. Either something is wrong, or these are estimates for different populations, and hence not perfectly comparable.

- **Residential growth.** In this `PGE_gas_plan_data`, we can see a strong long-term decline in Residential consumption --- CY1990 was 65% higher than CY2011, so about -3% per year. That's a full order of magnitude larger than the 0.3%/yr we saw in `SFBA_gas_county_data`.

- **Commercial/Industrial consumption (MMthm).** The amount of natural gas consumed by the Commercial and Industrial sectors, as reported in `PGE_gas_plan_data`, is almost twice as high as that labeled "Non-Residential" in `SFBA_gas_county_data` (3 billion vs 1.5 billion therms, respectively). 

- **Commercial/Industrial growth.** Even though the two datasets don't cover the same non-residential population, compared to the "Non-Residential" consumption in `SFBA_gas_county_data`, the long-term growth in the Commercial and Industrial sectors here looks similarly flat --- a fraction of a percent per year.

```{block type="note"}
Recall that `PGE_gas_plan_data` comprises almost 5 billion therms, whereas `SFBA_gas_county_data` comprised only 3 billion. So, these differences might be reconcilable. There could be large populations of consumers covered by one that aren't covered by the other.
```

**This section was intended** to give you a sense of what you can do with third-party activity data. You can do very similar things, of course, with third-party emission data. The key is to have the expected variables in your dataset (`tput_qty` and `tput_unit`, and/or `ems_qty` and `ems_unit`). And, you can always fall back to the more generic `chart_annual_quantities()` and/or `chart_annual_growth()`, so long as your dataset has at least one variable ending in `_qty`. 
