# Changes in Methodology

```{r SFBA_VMT_data, include = FALSE}
SFBA_VMT_data <-
  read_csv(
    here::here(
      "data",
      "SFBA-VMT-data.csv")) %>%
  transmute(
    year = CY(calendar_year),
    vtype = vtype,
    vmt_unit = "VMT/yr",
    vmt_qty = round(1e9 * vmt_qty))
```

```{r chart_data, include = FALSE}
chart_data <-
  SFBA_VMT_data %>%
  mutate(
    vmt_basis = if_else(
      year < CY(2000),
      "EMFAC2011",
      "EMFAC2017"))
```

"Changes in methodology" can be shown visually. Here, we do this for VMT data. 
In our dataset, `vmt_basis` is either "EMFAC2011" (for years CY1990-1999) or "EMFAC2017" (for years CY2000 onward). We use `vmt_basis` as the variable to color by. This introduces a visual break.

```{r SFBA_VMT_data-chart}
#
# NOTE: here we're omitting the usual titling and captioning ... just so that we
# can see that `color = vmt_basis` is the key piece.
# 
chart_data %>%
  chart_annual_quantities_by(
    color = vmt_basis)
```

The visual break is perhaps uncomfortable, but it is meaningful. It means that we don't actually have an estimate of the year-over-year change between CY1999 and CY2000! The reader should not be able to see a downward-sloping segment between those two years, unless we have reason to believe that there actually was a decrease.

## Annotating Your Data with `*_basis`

This is assuming we already have our `chart_data` in tidy form. Suppose we have our data in this form, but without a `*_basis` column.

```{r SFBA_VMT_data-show}
show(head(SFBA_VMT_data))
```

The key component is the `vmt_basis` column. For emission data, this might be `ems_basis`; for throughput data, it might be `tput_basis`. If you don't already have this in your data, you can add it using a combination of (a) `mutate()` and (b) either `if_else()` or `case_when()`.

```{r SFBA_VMT_data-mutate-if_else}
#
# Here is an example using `if_else()`.
#
chart_data <-
  SFBA_VMT_data %>%
  mutate(
    vmt_basis = if_else(
      year < CY(2000),
      true = "EMFAC2011",
      false = "EMFAC2017",
      missing = "Unknown"))
```

```{r SFBA_VMT_data-mutate-case_when}
#
# Here is an example using `case_when()`.
#
chart_data <-
  SFBA_VMT_data %>%
  mutate(
    vmt_basis = case_when(
      year < CY(2000) ~ "EMFAC2011",
      year >= CY(2000) ~ "EMFAC2017",
      TRUE ~ "Unknown"))
```