## Calculating county fractions from raw growth profile data {#appendix-howto-county_fractions}

In the chapter on [Growth profiles], things ended with a look at "raw" county-specific growth profile data for BY2011 category #761 "Sanitary Sewers". And, we could see that the raw data looked a lot like population counts.

This means that we can double-check the BY2011 county fractions, stored in file `r DataBank:::DB_activity_tbl(BY(2011))`, against fractions that we compute right here using the raw growth profile data. 

### Sanitary Sewers example {-}

Let's try this with our "Sanitary Sewers" example (which was linked to growth profile `#657`).

```{r BY2011-DB_growth_profile_crosswalk-cat_761}
#
# Double-check the crosswalk. Category #761 "Sanitary Sewers" should be
# associated with growth profile #657 "Household Population".
# 
BY(2011) %>%
  DB_growth_profile_crosswalk() %>%
  filter_categories(
    "Sanitary Sewers" = 761) %>%
  ensure(
    .$gpf_id == 657)
```

```{r BY2011_raw_growth_profile_data-gpf_657-mutate_cnty_frac}
#
# Compute county fractions from the raw growth profile data.
#
# - Extract just the CY2011 values.
# - Create a new column `cnty_frac`.
#
BY2011_raw_growth_profile_data %>%
  filter(
    gpf_id == 657) %>%
  filter_years(
    CY(2011)) %>%
  mutate(
    cnty_frac = gf_qty / total(gf_qty))
```

In the new `cnty_frac` column, we can see percentages of `r BY2011_raw_growth_profile_data %>% filter(gpf_id == 657) %>% filter_years(CY(2011)) %>% mutate(cnty_frac = gf_qty / sum(gf_qty)) %>% pull(cnty_frac) %>% format_percentage(digits = 1) %>% str_and()`, respectively.

### County fractions pulled from DataBank {-}

Now, let's compare those to what's stored in DataBank as "published" (file `r DataBank:::DB_activity_tbl(BY(2011))`, accessed via `DB_county_fractions()`). 

```{r BY2011-DB_county_fractions-cat_761}
#
# Pull the "published" county fractions for BY2011 (stored in `t1325`).
#
BY(2011) %>%
  DB_county_fractions(
    verbose = TRUE) %>%
  filter(
    cat_id == 761)
```

We can see some mismatches, but there could be good reasons for them. This is just intended to show you how you can do the double-checking.
