## Changes in forecasts for category #283 {#appendix-case-cat_283}

You might have noticed that the throughput forecast, for category #283, doesn't follow the same trajectory as the NOx forecast.

This is not because NOx emission factors were projected to increase after CY2011. Here's what happened:

1. The BY2011 criteria-pollutant calculations were completed. 
2. The growth profile for this category, in DataBank, was updated.
3. The BY2011 GHG calculations were completed. 

Here are some basic exercises (and charts) that may shed some light.

```{r appendix-BY2011_annual_emission_data-283-NOx, message = FALSE}
#
# This is the growth in NOx emissions from BY2011 category #283, via the
# published inventory (`BY2011_annual_emission_data`).
# 
BY2011_annual_emission_data %>%
  filter_categories(
    283) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_growth(
    base_year = CY(2011),
    flag_years = CY(2020)) 
```

```{r appendix-growth_profile-283, message = FALSE}
#
# This is the growth profile assigned to category #283 via file `t1325`. 
# It's not consistent with the published emission forecast (above).
# 
BY2011_growth_profile_data %>%
  filter_categories(
    283) %>%
  chart_annual_growth(
    base_year = CY(2011),
    flag_years = CY(2020)) 
```

```{r appendix-BY2011_area_source_projection_data-283-NOx, message = FALSE}
#
# This is the growth in emissions, via the reproduced inventory. It's consistent
# with the assigned growth profile, but not with the published forecast.
# 
BY2011_area_source_projection_data %>%
  filter_categories(
    283) %>%
  filter_pollutants(
    "NOx") %>%
  chart_annual_growth(
    qty_var = "ems_qty",
    base_year = CY(2011),
    flag_years = CY(2020)) 
```

```{r appendix-BY2011_area_source_throughput_data-283, message = FALSE}
#
# This is the growth in throughputs, via the reproduced inventory. It's also
# consistent with the assigned growth profile, and with the reproduced
# emissions, but not with the published forecast.
# 
BY2011_area_source_throughput_data %>%
  filter_categories(
    283) %>%
  chart_annual_growth(
    qty_var = "tput_qty",  # changed from "ems_qty" to "tput_qty"
    base_year = CY(2011),
    flag_years = CY(2020)) 
```

If it still doesn't make sense, please feel free to raise a question at one of our meetings!
